---
title: "Functional group analysis"
output: html_notebook
---

# Set up and data read in

```{r, warning=FALSE}
library(MetBrewer)
library(patchwork)
library(gggenes)
library(tidyverse)


# genomes and accessions
metadata <- read.csv(file.path("part2_functionalgroups", "metadata.csv"), header = TRUE)

# plotting

```

# cphA 


```{r}

genegrouper_path <- file.path("part2_functionalgroups/genegrouper/cpha")

genegroup_reps <- read.csv(file.path(genegrouper_path, "internal_data", "rtable_region_representatives.csv"), header = TRUE)
all_data <- read.csv(file.path(genegrouper_path, "group_regions.csv"))

```

## Group summary
Which bacteria is in each group? note that G-1 is every genome that didn't make a clean cluster

Note that only Dechloromonas phosphorivorans had one genome where cphA wasnt detected
```{r}

all_data %>% mutate(group = group + 1) %>%
  group_by(group) %>% summarise(num_unique = n_distinct(assembly_id))

metadata <- metadata %>% group_by(organism) %>% 
  mutate(num_species = n_distinct(assembly_id))

group_summary <- all_data %>% mutate(group = group + 1) %>%
  group_by(group) %>% reframe(assembly_id = unique(assembly_id)) %>% # get unique assembly ids by group
  ungroup() %>% left_join(metadata, by = "assembly_id") %>% # join to metadata 
  group_by(organism) %>% mutate(num_in_grp = n_distinct(assembly_id)) %>% # count number of species in each group
  mutate(perc_in_grp = num_in_grp / num_species) # get percent in each group


group_map_full <- 
ggplot(data = group_summary, aes(x = group.x, y = organism)) +
  geom_tile(color = "black", fill = "thistle") + 
  scale_y_discrete(limits=rev) +
  labs(x = "Group", y = "Organism", title = "") + 
  theme_classic() +
  theme(axis.line = element_blank(), axis.text=element_text(size=9))

ggsave(file.path("part2_functionalgroups/results", "group_map_full.tiff"), 
       height = 6, width = 5, units = "in", dpi = 300)

group_map <- 
  group_summary %>% filter(group.x != 0 ) %>%
  mutate(group.x = as.factor(group.x)) %>%
  ggplot(data = ., aes(x = group.x, y = organism)) +
  geom_tile(color = "black", fill = "thistle") + 
  scale_y_discrete(limits=rev) +
  labs(x = "Group", y = "Organism", title = "") + 
  theme_classic() +
  theme(axis.line = element_blank(), axis.text=element_text(size=9))

group_map

ggsave(file.path("part2_functionalgroups/results", "group_map.tiff"), 
       height = 6, width = 5, units = "in", dpi = 300)


```




## Gene maps of cphA clusters
Mapping gene clusters around cphA from genegrouper - genegrouper makes these plots automatically but I'm doing some customization 

```{r, fig.height = 8, fig.width = 8}

dfs <- genegroup_reps %>% filter(dbscan_label != -1) %>% 
  mutate(plt_grp = dbscan_label + 1) %>%
  mutate(refseq_gene = ifelse(refseq_product == "cyanophycinase", "cphB", refseq_gene)) %>%
  group_by(plt_grp) %>% group_split()

# apply ggplot function and write to list
gg_list <- lapply(dfs, function(x) {
  plt_title <- x[["plt_grp"]][1]
  numrow <- nrow(x) 
  ggplot(x, aes(xmin = cds_start, xmax = cds_end, y = 0, 
                           forward = (strand+1), fill = refseq_product, label = refseq_gene)) +
    geom_gene_arrow() + 
    geom_gene_label() +
    geom_text(aes(x = cds_end - ((cds_end - cds_start)/2), 
                  y = 0, label = refseq_product), angle = 40, vjust = -1.5, hjust = 0, size = 3) + 
    scale_x_continuous(expand = expansion(mult = c(0 , 0.2))) + 
    scale_y_continuous(limits = c(-0.01, 0.2), breaks = NULL) + 
    scale_fill_manual(values = met.brewer("Signac", n = numrow)) + 
    labs(x = "", y = "", title = paste("Group", plt_title)) + 
    theme_genes() +
    theme(legend.position = "none")
})

gene_clusters <- gg_list[[1]] / gg_list[[2]]

gene_clusters

ggsave(plot = gene_clusters, file.path("part2_functionalgroups/results", "gene_map_bygroup.tiff"),
       height = 4, width = 8, units = "in", dpi = 300)


```

```{r, fig.height = 6, fig.width = 12}

(group_map + labs(tags = "A") | ((gg_list[[1]] + labs(tags = "B")) / gg_list[[2]])) + 
  theme(plot.margin = margin(0,0,0,0, "cm")) + plot_layout(widths = c(1, 20)) 

ggsave(file.path("part2_functionalgroups/results", "gene_map_combo.tiff"),
       height = 5, width = 12, units = "in", dpi = 300)

```





## Summary stats by functional group
```{r}

matches <- unique(all_data[["assembly_id"]])
metadata <- metadata %>% mutate(cpha = ifelse(assembly_id %in% matches, TRUE, FALSE)) %>% 
  filter(!group %in% c("OHO", "CYANO"))

write.csv(metadata, file.path("part2_functionalgroups/table_S1.csv"), row.names = FALSE)


```



```{r}

metadata %>% filter(group == "AOB" & genus == "nitrosomonas")


```


```{r}

group_summary <- metadata %>% 
  group_by(group) %>% summarise(tot = n(), tot_yes = sum(cpha==TRUE))

group_summary
```



